---
title: "General Social Survey"
author: "Steven Garcia"
date: "Year 2020"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$ There is a difference between gun ownership and the opinion of the death penalty. 

$H_A$ There is a relationship between gun ownership and the opinion of the death penalty.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables are categorical, so we use the cat~cat as the helper file.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")

```

This graph shows that those who own a gun are more likely to favor the death penalty. Those opposed to death penalty are less likely to own a gun.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~OWNGUN + CAPPUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

Among those who own a gun, 75% favor capitol punishment and 25% oppose. Those who don't own a gun 57% favor capitol punishment compared to 42% oppose. Those who own guns are more likely to favor capitol punishmen compared to those who oppose. 

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test says the p-value is 0.02022, less that 0.05, so we reject the null hypothesis.

The fisher exact test p-value is 0.01651, less than 0.05, so we conclude that there is a relationship between gun ownership and opinion on capitol punishment. People who own a gun are on average 2.3 times more likely to favor capitol punishment. 

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$ There is a relationship between years of education and the number of voters in 2012. 

$H_A$ There is a relationship between years education and the the number of voters in 2012.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The response variable is numerical for education. The predictor variable is voters in 2012 with a categorical response. The helper file used will be numerical~cat.rmd.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

This graph shows that those who voted have a higher education level, compared to those with a lower level of education did not vote.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```

According to the numerical result, group that voted had a mean of 15.72, compared to the non voters had an average mean of 13.42.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

The p-value is 1.248e, which is less than 0.05, so we reject the null hypothesis. The data indicates that there is difference between the level of education and the relation to voters.  

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  

$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The response variable is numerical for the average score. The predictor variable is degree with a categorical response. The helper file used will be numerical~cat.rmd. 

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

I used chatGPT to get R code for plotting.

```{r}
# Box plot to visualize response vs. predictor
ggplot(data=GSSdata, aes( x = DEGREE, y = WORDSUM)) +
  geom_boxplot() +
  xlab("category") +
  ylab("WORDSUM") +
  ggtitle("Box Plot of Response variable by category")


```

The box plot indicates that the average score for the WORDSUM is the same for junior college and bachelor degrees. Graduate has the highest average score compared to the rest. Compared to high school and LT high schools with the lowest average scores. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE, data=GSSdata)
```

According to numerical result, the graduate degree category had the highest scoring mean compared to the rest. Following close behind is the bachelor and junior college degree graduate that had an average mean of 9. Where the high school and LT high school.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSSdata)
summary.aov(model1)
```

This shows the P value is less than 0.05, so we reject the null hypothesis. This indicates that there is a difference in WORDSUM tests compared to the level of degree. 

```{r}
plot(TukeyHSD(model1)) 
print(TukeyHSD(model1))
```

We do have evidence that there is a difference in levels of degree and the WORDSUM score. The biggest difference is the higher the education is directly related to higher WORDSUM score. If the line doesn't intersect the 0 line interval, that means there is no difference in test scores. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Is there difference between the average score on the WORDSUM test and level of schooling DEGREE?  After screening the information, the data indicates that there is a difference between the education level and scores for WORDSUM test. The graphical result show the higher educated individuals tested higher and the P value is below 0.05 and we then rejected the null hypothesis. The result is that there is a difference between the college level degree and the the test scores for WORDSUM.
